\vspace{-1cm} Model simplicity breeds contempt: using simple models to answer basic questions on species' distributions

### 2021-02-04 manuscript/manuscript.tex
Fixing convergence problem and figures

\begin{document}
\maketitle
\linenumbers
\section*{Introduction}
{\color{gray}
These are just a bunch of thoughts that I had on how to frame the introduction for the manuscript. I want to say something along the lines: \textit{we know a lot about the factors that could theoretically influence species' distributions, and a rapidly growing body of research have been primarily focused on trying to untangle some of such biotic and abiotic predictors---with an increasing effort placed in improving the predictive power of statistical models. However, much less is known about how species' distributions compare to each other. Here, we use a conceptually more conservative approach to instead understand and compare basic aspects regarding the shape of species' distribution along environmental gradients.}}
%\citet{austin2002spatial} writes "\textit{there are three major components in any framework for statistical modelling in plant ecology. There needs to be an ecological model, a data model, and a statistical model. The ecological model consists of the ecological knowledge and theory to be used or tested in the study. The data model consists of the decisions made regarding how the data are collected and how the data will be measured or estimated. The statistical model involves the choice of statistical method, error function and significance tests. Each model interacts in both obvious and subtle ways with the other models to determine the success of any statistical modelling exercise.}" While a lot of work has been developed for all three components outlined by \citet{austin2002spatial}, increasing emphasis has been put into advancing the data and statistical models. Indeed,

@@

>>>>

\begin{document}
\maketitle
\linenumbers
\section*{Introduction}
Accounting for expert knowledge on species' environmental preferences to understand general distribution patterns.

The use of species distribution models has grown a lot. These models try to estimate species' realized niches using several covariates, including environmental variables (ref), species ecological traits' (ref) and phylogenetic relations (ref). Recent work on these modelling approaches has increasingly focused on estimating (ref) and accounting for (ref) biotic factors, such as competitive or facilitative relationships. The idea is that by understanding how all such factors shape species' distribution we will gain a mechanistic understanding on how these distributions are established and change over time. Unfortunately, while some of this approaches will certainly increase the predictive performance of distribution models (ref), the nature of some of the estimates have been shown to theoretically have some mishaps (ref).

{\color{gray}
These are just a bunch of thoughts that I had on how to frame the introduction for the manuscript. I want to say something along the lines: \textit{we know a lot about the factors that could theoretically influence species' distributions, and a rapidly growing body of research have been primarily focused on trying to untangle some of such biotic and abiotic predictors---with an increasing effort placed in improving the predictive power of statistical models. However, much less is known about how species' distributions compare to each other. Here, we use a conceptually more conservative approach to instead understand and compare basic aspects regarding the shape of species' distribution along environmental gradients.}}
%\citet{austin2002spatial} writes "\textit{there are three major components in any framework for statistical modelling in plant ecology. There needs to be an ecological model, a data model, and a statistical model. The ecological model consists of the ecological knowledge and theory to be used or tested in the study. The data model consists of the decisions made regarding how the data are collected and how the data will be measured or estimated. The statistical model involves the choice of statistical method, error function and significance tests. Each model interacts in both obvious and subtle ways with the other models to determine the success of any statistical modelling exercise.}" While a lot of work has been developed for all three components outlined by \citet{austin2002spatial}, increasing emphasis has been put into advancing the data and statistical models. Indeed,

@@

%%%%%%%%%%%%%%%%%%%%%%%
% Paragraph on Rapoport's rule
%%%%%%%%%%%%%%%%%%%%%%%
% McCain
%Understanding the factors that determine species range size dis- tributions is becoming increasingly urgent as more climate change assessments document the heightened risk for small- ranged species (e.g. Channell & Lomolino, 2000; Sekercioglu et al., 2008; La Sorte & Jetz, 2010; McCain & Colwell, 2011).
%Rapoport’s rule is the positive relationship of species range sizes with increasing latitude (Stevens, 1989), elevation (Stevens, 1992) or water depth (Stevens, 1996). The latitudinal Rapoport’s rule (LRR) is the most examined in the literature, including testing empirical datasets (e.g. Rohde et al., 1993; Roy et al., 1994; Blackburn & Gaston, 1996; Lyons & Willig, 1997; Ruggiero & Lawton, 1998; Reed, 2003; Arita et al., 2005; Hausdorf, 2006), simulation modelling of the expected patterns (Colwell & Hurtt, 1994; Taylor & Gaines, 1999; Case & Taper, 2000; Arita, 2005; Stauffer & Rohde, 2006; Šizling et al., 2009), and theoretical and empirical evidence for the proposed mechanisms (e.g. Rohde, 1992; Kerr, 1999; Taylor & Gaines, 1999; Gaston & Chown, 1999a; Addo-Bediako et al., 2000; Parmesan et al., 2005). The various reviews of the LRR to date suggest that the overall support is weak (e.g. Rohde, 1996; Gaston et al., 1998; Rohde, 1999; Gaston & Chown, 1999b; Ribas & Schoereder, 2006), prin- cipally due to the high degree of variability in the fit to predic- tions (e.g. support: Blackburn & Gaston, 1996; Lyons & Willig, 1997; Price et al., 1997; Arita et al., 2005; no support: Rohde et al., 1993; Roy et al., 1994; Ruggiero & Lawton, 1998; Reed, 2003).
%The elevational Rapoport’s rule (ERR) has received less examination in the literature than the LRR, but there have also been empirical tests of its predictions (e.g. Patterson et al., 1996; Pleguezuelos & Villafranca, 1997; Price et al., 1997; Rahbek, 1997; Fleishman et al., 1998; Ruggiero & Lawton, 1998; Nathan & Werner, 1999; Sanders, 2002; Fu et al., 2004; Chatzaki et al., 2005; Almeida-Neto et al., 2006; Bhattarai & Vetaas, 2006; Hausdorf, 2006) and underlying theory (e.g. Rohde, 1996; Fleishman et al., 1998; Gaston & Chown, 1999a; Hausdorf, 2006; Ribas & Schoereder, 2006). In the ERR, like the LRR, there is a high degree of variability in support from supportive (e.g. Pat- terson et al., 1996; Pleguezuelos & Villafranca, 1997; Price et al., 1997; Rahbek, 1997; Fleishman et al., 1998; Sanders, 2002; Chatzaki et al., 2005; Almeida-Neto et al., 2006; Hausdorf, 2006; Ribas & Schoereder, 2006) to little or no support (e.g. Patterson et al., 1996; Price et al., 1997; Rahbek, 1997; Ruggiero & Lawton, 1998; Nathan & Werner, 1999; Fu et al., 2004; Bhattarai & Vetaas, 2006; Hausdorf, 2006; Ribas & Schoereder, 2006). Similarly, the processes proposed to underlie the ERR (Stevens, 1992) also show a variable amount of theoretical and empirical support (Patterson et al., 1996; Rohde, 1996; Rahbek, 1997; Fleishman et al., 1998; Gaston & Chown, 1999a; Almeida-Neto et al., 2006; Hausdorf, 2006; Colwell, 2011).
Scarce data and little to no attempt to account for uncertainty in the predictions. Similar to rapoport's rule, we can also ask other questions regarding general geographical patterns of species distributions.
